\pkg{casebase}: An Alternative Framework For Survival Analysis

### 2019-12-08 analysis/paper/paper.Rmd

---
documentclass: jss
author:
  - name: Sahir Bhatnagar *
    affiliation: McGill Univeristy
    address: >
      1020 Pine Avenue West
      Montreal, QC, Canada H3A 1A2
    email: \email{sahir.bhatnagar@mail.mcgill.ca}
    url: http://sahirbhatnagar.com/
  - name: Maxime Turgeon *
    affiliation: 'University of Manitoba \AND'
    address: >
      186 Dysart Road
      Winnipeg, MB, Canada R3T 2N2
    email: \email{max.turgeon@umanitoba.ca}
    url: https://maxturgeon.ca/
  - name: Jesse Islam 
    affiliation: McGill University
    address: >
      1020 Pine Avenue West
      Montreal, QC, Canada H3A 1A2
    email: \email{jesse.islam@mail.mcgill.ca}
  - name: James Hanley
    affiliation: McGill Univeristy
    address: >
      1020 Pine Avenue West
      Montreal, QC, Canada H3A 1A2
    email: \email{james.hanley@mcgill.ca}
    url: http://www.medicine.mcgill.ca/epidemiology/hanley/
  - name: Olli Saarela
    affiliation: University of Toronto
    address: >
      Dalla Lana School of Public Health, 
      155 College Street, 6th floor, 
      Toronto, Ontario 
      M5T 3M7, Canada
    email: \email{olli.saarela@utoronto.ca}
    url: http://individual.utoronto.ca/osaarela/
title:
  formatted: "\\pkg{casebase}: An Alternative Framework For Survival Analysis"
  # If you use tex in the formatted title, also supply version without
  plain:     "casebase: An Alternative Framework For Survival Analysis"
  # For running headers, if needed
  short:     "\\pkg{casebase}: An Alternative Framework For Survival Analysis"
abstract: >
  The abstract of the article. * joint co-authors
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
  \usepackage{longtable}
output: 
  rticles::jss_article:
    number_sections: TRUE     #added argument option 
    citation_package: "natbib"  #All my citations use biblatex, not natbib. 
biblio-style: jss      #Listed to use in JSS Instructions for Authors, but not in template by default. 
bibliography: references.bib  #Also not included in template by default. 
fig_width: 7
fig_height: 6
fig_caption: true
---

# Code formatting {.unnumbered}

Don't use markdown, instead use the more precise latex commands:

* \proglang{Java}
* \pkg{plyr}
* \code{print("abc")}

@@

>>>>

---
documentclass: jss
author:
  - name: Sahir Bhatnagar *
    affiliation: McGill University
    address: >
      1020 Pine Avenue West
      Montreal, QC, Canada H3A 1A2
    email: \email{sahir.bhatnagar@mail.mcgill.ca}
    url: http://sahirbhatnagar.com/
  - name: Maxime Turgeon *
    affiliation: 'University of Manitoba \AND'
    address: >
      186 Dysart Road
      Winnipeg, MB, Canada R3T 2N2
    email: \email{max.turgeon@umanitoba.ca}
    url: https://maxturgeon.ca/
  - name: Jesse Islam 
    affiliation: McGill University
    address: >
      1020 Pine Avenue West
      Montreal, QC, Canada H3A 1A2
    email: \email{jesse.islam@mail.mcgill.ca}
  - name: James Hanley
    affiliation: McGill University
    address: >
      1020 Pine Avenue West
      Montreal, QC, Canada H3A 1A2
    email: \email{james.hanley@mcgill.ca}
    url: http://www.medicine.mcgill.ca/epidemiology/hanley/
  - name: Olli Saarela
    affiliation: University of Toronto
    address: >
      Dalla Lana School of Public Health, 
      155 College Street, 6th floor, 
      Toronto, Ontario 
      M5T 3M7, Canada
    email: \email{olli.saarela@utoronto.ca}
    url: http://individual.utoronto.ca/osaarela/
title:
  formatted: "\\pkg{casebase}: An Alternative Framework For Survival Analysis"
  # If you use tex in the formatted title, also supply version without
  plain:     "casebase: An Alternative Framework For Survival Analysis"
  # For running headers, if needed
  short:     "\\pkg{casebase}: An Alternative Framework For Survival Analysis"
abstract: >
  The abstract of the article. * joint co-authors
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
  \usepackage{longtable}
output: 
  rticles::jss_article:
    number_sections: TRUE     #added argument option 
    citation_package: "natbib"  #All my citations use biblatex, not natbib. 
biblio-style: jss      #Listed to use in JSS Instructions for Authors, but not in template by default. 
bibliography: references.bib  #Also not included in template by default. 
fig_width: 7
fig_height: 6
fig_caption: true
---

<!-- # Code formatting {.unnumbered} -->

<!-- Don't use markdown, instead use the more precise latex commands: -->

<!-- * \proglang{Java} -->
<!-- * \pkg{plyr} -->
<!-- * \code{print("abc")} -->

@@

# Case study 4--Stanford Heart Transplant Data

Although the previous case studies provide a broad overview of the capabilities of \pkg{casebase}, they all have two properties in common:
\begin{itemize}
  \item All covariates are fixed, i.e. they do not change over time;
  \item The estimated hazard functions satisfy the "proportional hazard" assumption.
\end{itemize}

Case-base sampling has been used in the literature to study vaccination safety, where the exposure period was defined as the week following vaccination [@saarela2015case]. Hence, the main covariate of interest, i.e. exposure to the vaccine, was changing over time. In this context, case-base sampling offers an efficient alternative to nested case-control designs or self-matching.

The purpose of the next case study is to show how \pkg{casebase} can also be used for survival analysis problems that do not share the properties above. To this end, we will use the Stanford Heart Transplant Data [@clark1971cardiac,@crowley1977covariance]; we will use the version available in the \pkg{survival} package.

Recall the setting of this study: patients were admitted to the Stanford program after meeting with their physician and determining that they were unlikely to respond to other forms of treatment. After enrollment, the program searches for a suitable donor for the patient, which can take anywhere between a few days to almost a year. We are interested in the effect of a heart transplant on survival; therefore, the patient is considered exposed only after the transplant has occured.

As before, we can look at the population-time plot for a graphical summary of the event incidence. As we can see, most events occur early during the follow-up period, and therefore we do not expect the hazard to be constant.

```{r stanford-poptime, eval = eval_cs4}
library(survival)
library(casebase)

stanford_popTime <- popTime(jasa, time = "futime", 
                            event = "fustat")
plot(stanford_popTime)
```

Since the exposure is time-dependent, we need to manually define the exposure variable *after* case-base sampling and *before* fitting the hazard function. For this reason, we will use the `sampleCaseBase` function directly.

```{r, eval = eval_cs4}
library(tidyverse)
library(lubridate)

cb_data <- sampleCaseBase(jasa, time = "futime", 
                          event = "fustat", ratio = 10)
```

Next, we will compute the number of days from acceptance into the program to transplant, and we use this variable to determine whether each population-moment is exposed or not.

```{r, eval = eval_cs4}
# Define exposure variable
cb_data <- mutate(cb_data,
                  txtime = time_length(accept.dt %--% tx.date, 
                                       unit = "days"),
                  exposure = case_when(
                    is.na(txtime) ~ 0L,
                    txtime > futime ~ 0L,
                    txtime <= futime ~ 1L
                  ))
```

Finally, we can fit the hazard using various linear predictors.

```{r, eval = eval_cs4, warning = FALSE}
library(splines)
# Fit several models
fit1 <- fitSmoothHazard(fustat ~ exposure,
                        data = cb_data, time = "futime")
fit2 <- fitSmoothHazard(fustat ~ exposure + futime,
                        data = cb_data, time = "futime")
fit3 <- fitSmoothHazard(fustat ~ exposure + bs(futime),
                        data = cb_data, time = "futime")
fit4 <- fitSmoothHazard(fustat ~ exposure*bs(futime),
                        data = cb_data, time = "futime")
```

Note that the fourth model (i.e. `fit4`) includes an interaction term between exposure and follow-up time. In other words, this model no longer exhibit proportional hazards. The evidence of non-proportionality of hazards in the Stanford Heart Transplant data has been widely discussed [@arjas1988graphical].

We can then compare the goodness of fit of these four models using the Akaike Information Criterion (AIC).

```{r, eval = eval_cs4}
# Compute AIC
c("Model1" = AIC(fit1),
  "Model2" = AIC(fit2),
  "Model3" = AIC(fit3),
  "Model4" = AIC(fit4))
```

As we can, the best fit is the fourth model. By visualizing the hazard functions for both exposed and unexposed individuals, we can more clearly see how the hazards are no longer proportional.

```{r stanford-hazard, eval = eval_cs4}
# Compute hazards---
# First, create a list of time points for both exposure status
hazard_data <- expand.grid(exposure = c(0, 1),
                           futime = seq(0, 1000,
                                        length.out = 100))
# Set the offset to zero
hazard_data$offset <- 0 
# Use predict to get the fitted values, and exponentiate to 
# transform to the right scale
hazard_data$hazard = exp(predict(fit4, newdata = hazard_data,
                                 type = "link"))
# Add labels for plots
hazard_data$Status = factor(hazard_data$exposure,
                            labels = c("NoTrans", "Trans"))

ggplot(hazard_data, aes(futime, hazard, colour = Status)) +
    geom_line() +
    theme_minimal() +
    theme(legend.position = 'top') +
    ylab('Hazard') + xlab('Follow-up time')
```

The non-proportionality seems to be more pronounced at the beginning of follow-up than the end. Finally, we can turn these estimate of the hazard function into estimates of the cumulative incidence functions.

```{r stanford-risk, eval = eval_cs4}
# Compute absolute risk curves
newdata <- data.frame(exposure = c(0, 1))
absrisk <- absoluteRisk(fit4, newdata = newdata, 
                        time = seq(0, 1000, length.out = 100))

colnames(absrisk) <- c("Time", "NoTrans", "Trans")

# Rearrange the data
absrisk <- gather(as.data.frame(absrisk),
                  "Status", "Risk", -Time)

ggplot(absrisk, aes(Time, Risk, colour = Status)) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = 'top') +
  expand_limits(y = 0.5) +
  xlab('Follow-up time') + ylab('Cum. Incidence')
```

Note that we can easily adapt the code above to the situation where a patient receives a heart transplant at a point in time of interest, for example after 30 days.

```{r stanford-1y-haz, eval = eval_cs4}
# Compute hazards---
# First, create a list of time points for both exposure status
one_yr_haz <- data.frame(futime = seq(0, 365,
                                      length.out = 100))
one_yr_haz <- mutate(one_yr_haz,
                     offset = 0,
                     exposure = if_else(futime < 30, 0, 1))
one_yr_haz$hazard = exp(predict(fit4, newdata = one_yr_haz,
                                type = "link"))
ggplot(one_yr_haz, aes(futime, hazard)) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = 'top') +
  ylab('Hazard') + xlab('Follow-up time') +
  geom_vline(xintercept = 30, linetype = 'dashed')
```

We can then compare the 1-year mortality risk without transplant and with transplant at 30 days.

```{r stanford-1y-risk, eval = eval_cs4}
absoluteRisk(fit4, newdata = data.frame(exposure = 0),
             time = 365)

# Use the trapezoidal rule to estimate the cumulative hazard
min_hazard <- min(one_yr_haz$hazard)
max_hazard <- max(one_yr_haz$hazard)
increment <- 365/99
cumhaz_est <- 0.5*increment*(sum(2*one_yr_haz$hazard) - min_hazard - max_hazard)
1 - exp(-cumhaz_est)
```

As we can see, the risk estimate at 1-year is about 30% lower if the patient receives a heart transplant at 30 days.

### 2019-12-08 analysis/paper/paper.tex

\Address{
    Sahir Bhatnagar *\\
  McGill University\\
  1020 Pine Avenue West Montreal, QC, Canada H3A 1A2\\
  E-mail: \email{sahir.bhatnagar@mail.mcgill.ca}\\
  URL: \url{http://sahirbhatnagar.com/}\\~\\
      Maxime Turgeon *\\
  University of Manitoba\\
  186 Dysart Road Winnipeg, MB, Canada R3T 2N2\\
  E-mail: \email{max.turgeon@umanitoba.ca}\\
  URL: \url{http://maxturgeon.ca/}\\~\\
      Jesse Islam\\
  McGill University\\
  1020 Pine Avenue West Montreal, QC, Canada H3A 1A2\\
  E-mail: \email{jesse.islam@mail.mcgill.ca}\\

@@

Note that the fourth model (i.e. \texttt{fit4}) includes an interaction
term between exposure and follow-up time. In other words, this model no
longer exhibit proportional hazards. The evidence of non-proportionality
of hazards in the Stanford Heart Transplant data has been widely
discussed \citep{arjas1988graphical}.

@@

This report was generated on 2019-12-06 10:11:23 using the following
computational environment and dependencies:

@@

#> - Session info ---------------------------------------------------------------

@@

#>  date     2019-12-06                  

@@

#> - Packages -------------------------------------------------------------------
#>  package     * version    date       lib
#>  assertthat    0.2.1      2019-03-21 [1]
#>  backports     1.1.5      2019-10-02 [1]
#>  broom         0.5.2      2019-04-07 [1]
#>  callr         3.3.2      2019-09-22 [1]
#>  casebase    * 0.2.1.9001 2019-08-13 [1]
#>  cellranger    1.1.0      2016-07-27 [1]
#>  cli           1.1.0      2019-03-19 [1]
#>  colorspace    1.4-1      2019-03-18 [1]
#>  crayon        1.3.4      2017-09-16 [1]
#>  data.table    1.12.6     2019-10-18 [1]
#>  DBI           1.0.0      2018-05-02 [1]
#>  dbplyr        1.4.2      2019-06-17 [1]
#>  desc          1.2.0      2018-05-01 [1]
#>  devtools      2.2.1      2019-09-24 [1]
#>  digest        0.6.23     2019-11-23 [1]
#>  dplyr       * 0.8.3      2019-07-04 [1]
#>  ellipsis      0.3.0      2019-09-20 [1]
#>  evaluate      0.14       2019-05-28 [1]
#>  farver        2.0.1      2019-11-13 [1]
#>  forcats     * 0.4.0      2019-02-17 [1]
#>  fs            1.3.1      2019-05-06 [1]
#>  generics      0.0.2      2018-11-29 [1]
#>  ggplot2     * 3.2.1      2019-08-10 [1]
#>  glue          1.3.1      2019-03-12 [1]
#>  gtable        0.3.0      2019-03-25 [1]
#>  haven         2.2.0      2019-11-08 [1]
#>  hms           0.5.2      2019-10-30 [1]
#>  htmltools     0.4.0      2019-10-04 [1]
#>  httr          1.4.1      2019-08-05 [1]
#>  jsonlite      1.6        2018-12-07 [1]
#>  knitr         1.26       2019-11-12 [1]
#>  labeling      0.3        2014-08-23 [1]
#>  lattice       0.20-38    2018-11-04 [4]
#>  lazyeval      0.2.2      2019-03-15 [1]
#>  lifecycle     0.1.0      2019-08-01 [1]
#>  lubridate   * 1.7.4      2018-04-11 [1]
#>  magrittr    * 1.5        2014-11-22 [1]
#>  Matrix        1.2-17     2019-03-22 [4]
#>  memoise       1.1.0      2017-04-21 [1]
#>  mgcv          1.8-31     2019-11-09 [4]
#>  modelr        0.1.5      2019-08-08 [1]
#>  munsell       0.5.0      2018-06-12 [1]
#>  nlme          3.1-142    2019-11-07 [4]
#>  pillar        1.4.2      2019-06-29 [1]
#>  pkgbuild      1.0.6      2019-10-09 [1]
#>  pkgconfig     2.0.3      2019-09-22 [1]
#>  pkgload       1.0.2      2018-10-29 [1]
#>  prettyunits   1.0.2      2015-07-13 [1]
#>  processx      3.4.1      2019-07-18 [1]
#>  ps            1.3.0      2018-12-21 [1]
#>  purrr       * 0.3.3      2019-10-18 [1]
#>  R6            2.4.1      2019-11-12 [1]
#>  Rcpp          1.0.3      2019-11-08 [1]
#>  readr       * 1.3.1      2018-12-21 [1]
#>  readxl        1.3.1      2019-03-13 [1]
#>  remotes       2.1.0      2019-06-24 [1]
#>  reprex        0.3.0      2019-05-16 [1]
#>  rlang         0.4.2      2019-11-23 [1]
#>  rmarkdown     1.18       2019-11-27 [1]
#>  rprojroot     1.3-2      2018-01-03 [1]
#>  rstudioapi    0.10       2019-03-19 [1]
#>  rticles       0.12       2019-11-12 [1]
#>  rvest         0.3.5      2019-11-08 [1]
#>  scales        1.1.0      2019-11-18 [1]
#>  sessioninfo   1.1.1      2018-11-05 [1]
#>  stringi       1.4.3      2019-03-12 [1]
#>  stringr     * 1.4.0      2019-02-10 [1]
#>  survival    * 2.44-1.1   2019-04-01 [4]
#>  testthat      2.3.1      2019-12-01 [1]
#>  tibble      * 2.1.3      2019-06-06 [1]
#>  tidyr       * 1.0.0      2019-09-11 [1]
#>  tidyselect    0.2.5      2018-10-11 [1]
#>  tidyverse   * 1.3.0      2019-11-21 [1]
#>  usethis       1.5.1      2019-07-04 [1]
#>  vctrs         0.2.0      2019-07-05 [1]
#>  VGAM          1.1-2      2019-11-21 [1]
#>  withr         2.1.2      2018-03-15 [1]
#>  xfun          0.11       2019-11-12 [1]
#>  xml2          1.2.2      2019-08-09 [1]
#>  yaml          2.2.0      2018-07-25 [1]
#>  zeallot       0.1.0      2018-01-28 [1]
#>  source                                  
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  Github (sahirbhatnagar/casebase@aecb816)
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.5.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          
#>  CRAN (R 3.6.1)                          

@@

#> [1] /home/turgeonm/R/x86_64-pc-linux-gnu-library/3.6

@@

\begin{CodeOutput}
#> Local:    review-max /home/turgeonm/Documents/git_repos/cbpaper
#> Head:     [d0d7b8a] 2019-12-05: add analysis of stanford data
\end{CodeOutput}
\end{CodeChunk}

>>>>

\Address{
    Sahir Bhatnagar *\\
  McGill University\\
  1020 Pine Avenue West Montreal, QC, Canada H3A 1A2\\
  E-mail: \email{sahir.bhatnagar@mail.mcgill.ca}\\
  URL: \url{http://sahirbhatnagar.com/}\\~\\
      Maxime Turgeon *\\
  University of Manitoba\\
  186 Dysart Road Winnipeg, MB, Canada R3T 2N2\\
  E-mail: \email{max.turgeon@umanitoba.ca}\\
  URL: \url{https://maxturgeon.ca/}\\~\\
      Jesse Islam\\
  McGill University\\
  1020 Pine Avenue West Montreal, QC, Canada H3A 1A2\\
  E-mail: \email{jesse.islam@mail.mcgill.ca}\\

@@

Note that the fourth model (i.e.~\texttt{fit4}) includes an interaction
term between exposure and follow-up time. In other words, this model no
longer exhibit proportional hazards. The evidence of non-proportionality
of hazards in the Stanford Heart Transplant data has been widely
discussed \citep{arjas1988graphical}.

@@

This report was generated on 2019-12-08 11:49:58 using the following
computational environment and dependencies:

@@

#> - Session info ----------------------------------------------------------

@@

#>  date     2019-12-08                  

@@

#> - Packages --------------------------------------------------------------
#>  package     * version    date       lib source                          
#>  assertthat    0.2.1      2019-03-21 [1] CRAN (R 3.6.1)                  
#>  backports     1.1.5      2019-10-02 [1] CRAN (R 3.6.1)                  
#>  broom         0.5.2      2019-04-07 [1] CRAN (R 3.6.1)                  
#>  callr         3.3.1      2019-07-18 [1] CRAN (R 3.6.1)                  
#>  casebase    * 0.2.1.9001 2019-07-29 [1] local                           
#>  cellranger    1.1.0      2016-07-27 [1] CRAN (R 3.5.0)                  
#>  cli           1.1.0      2019-03-19 [1] CRAN (R 3.6.1)                  
#>  colorspace    1.4-1      2019-03-18 [1] CRAN (R 3.6.1)                  
#>  crayon        1.3.4      2017-09-16 [1] CRAN (R 3.5.0)                  
#>  data.table    1.12.2     2019-04-07 [1] CRAN (R 3.6.1)                  
#>  desc          1.2.0      2018-05-01 [1] CRAN (R 3.5.0)                  
#>  devtools      2.1.0      2019-07-06 [1] CRAN (R 3.6.1)                  
#>  digest        0.6.22     2019-10-21 [1] CRAN (R 3.6.1)                  
#>  dplyr       * 0.8.3      2019-07-04 [1] CRAN (R 3.6.1)                  
#>  ellipsis      0.3.0      2019-09-20 [1] CRAN (R 3.6.1)                  
#>  evaluate      0.14       2019-05-28 [1] CRAN (R 3.6.1)                  
#>  forcats     * 0.4.0      2019-02-17 [1] CRAN (R 3.6.1)                  
#>  fs            1.3.1      2019-05-06 [1] CRAN (R 3.6.1)                  
#>  generics      0.0.2      2018-11-29 [1] CRAN (R 3.6.1)                  
#>  ggplot2     * 3.2.1      2019-08-10 [1] CRAN (R 3.6.1)                  
#>  glue          1.3.1      2019-03-12 [1] CRAN (R 3.6.1)                  
#>  gtable        0.3.0      2019-03-25 [1] CRAN (R 3.6.1)                  
#>  haven         2.1.1      2019-07-04 [1] CRAN (R 3.6.1)                  
#>  hms           0.5.2      2019-10-30 [1] CRAN (R 3.6.1)                  
#>  htmltools     0.3.6      2017-04-28 [1] CRAN (R 3.5.0)                  
#>  httr          1.4.1      2019-08-05 [1] CRAN (R 3.6.1)                  
#>  jsonlite      1.6        2018-12-07 [1] CRAN (R 3.6.1)                  
#>  knitr         1.23       2019-05-18 [1] CRAN (R 3.6.1)                  
#>  labeling      0.3        2014-08-23 [1] CRAN (R 3.5.0)                  
#>  lattice       0.20-38    2018-11-04 [4] CRAN (R 3.5.1)                  
#>  lazyeval      0.2.2      2019-03-15 [1] CRAN (R 3.6.1)                  
#>  lifecycle     0.1.0      2019-08-01 [1] CRAN (R 3.6.1)                  
#>  lubridate   * 1.7.4      2018-04-11 [1] CRAN (R 3.5.0)                  
#>  magrittr    * 1.5        2014-11-22 [1] CRAN (R 3.5.0)                  
#>  Matrix        1.2-18     2019-11-27 [4] CRAN (R 3.6.1)                  
#>  memoise       1.1.0      2017-04-21 [1] CRAN (R 3.5.0)                  
#>  mgcv          1.8-24     2018-06-18 [1] CRAN (R 3.5.1)                  
#>  modelr        0.1.5      2019-08-08 [1] CRAN (R 3.6.1)                  
#>  munsell       0.5.0      2018-06-12 [1] CRAN (R 3.6.1)                  
#>  nlme          3.1-142    2019-11-07 [4] CRAN (R 3.6.1)                  
#>  pillar        1.4.2      2019-06-29 [1] CRAN (R 3.6.1)                  
#>  pkgbuild      1.0.3      2019-03-20 [1] CRAN (R 3.6.1)                  
#>  pkgconfig     2.0.3      2019-09-22 [1] CRAN (R 3.6.1)                  
#>  pkgload       1.0.2      2018-10-29 [1] CRAN (R 3.6.1)                  
#>  prettyunits   1.0.2      2015-07-13 [1] CRAN (R 3.5.2)                  
#>  processx      3.4.1      2019-07-18 [1] CRAN (R 3.6.1)                  
#>  ps            1.3.0      2018-12-21 [1] CRAN (R 3.6.1)                  
#>  purrr       * 0.3.3      2019-10-18 [1] CRAN (R 3.6.1)                  
#>  R6            2.4.1      2019-11-12 [1] CRAN (R 3.6.1)                  
#>  Rcpp          1.0.3      2019-11-08 [1] CRAN (R 3.6.1)                  
#>  readr       * 1.3.1      2018-12-21 [1] CRAN (R 3.6.1)                  
#>  readxl        1.3.1      2019-03-13 [1] CRAN (R 3.6.1)                  
#>  remotes       2.1.0      2019-06-24 [1] CRAN (R 3.6.1)                  
#>  rlang         0.4.1      2019-10-24 [1] CRAN (R 3.6.1)                  
#>  rmarkdown     1.16       2019-10-01 [1] CRAN (R 3.6.1)                  
#>  rprojroot     1.3-2      2018-01-03 [1] CRAN (R 3.5.0)                  
#>  rstudioapi    0.10       2019-03-19 [1] CRAN (R 3.6.1)                  
#>  rticles       0.9.1      2019-07-22 [1] Github (rstudio/rticles@8d56fc6)
#>  rvest         0.3.5      2019-11-08 [1] CRAN (R 3.6.1)                  
#>  scales        1.0.0      2018-08-09 [1] CRAN (R 3.6.1)                  
#>  sessioninfo   1.1.1      2018-11-05 [1] CRAN (R 3.6.1)                  
#>  stringi       1.4.3      2019-03-12 [1] CRAN (R 3.6.1)                  
#>  stringr     * 1.4.0      2019-02-10 [1] CRAN (R 3.6.1)                  
#>  survival    * 2.44-1.1   2019-04-01 [4] CRAN (R 3.6.1)                  
#>  testthat      2.2.0      2019-07-22 [1] CRAN (R 3.6.1)                  
#>  tibble      * 2.1.3      2019-06-06 [1] CRAN (R 3.6.1)                  
#>  tidyr       * 1.0.0      2019-09-11 [1] CRAN (R 3.6.1)                  
#>  tidyselect    0.2.5      2018-10-11 [1] CRAN (R 3.5.2)                  
#>  tidyverse   * 1.2.1      2017-11-14 [1] CRAN (R 3.5.0)                  
#>  usethis       1.5.1      2019-07-04 [1] CRAN (R 3.6.1)                  
#>  vctrs         0.2.0      2019-07-05 [1] CRAN (R 3.6.1)                  
#>  VGAM          1.1-1      2019-02-18 [1] CRAN (R 3.6.1)                  
#>  withr         2.1.2      2018-03-15 [1] CRAN (R 3.5.0)                  
#>  xfun          0.8        2019-06-25 [1] CRAN (R 3.6.1)                  
#>  xml2          1.2.2      2019-08-09 [1] CRAN (R 3.6.1)                  
#>  yaml          2.2.0      2018-07-25 [1] CRAN (R 3.6.1)                  
#>  zeallot       0.1.0      2018-01-28 [1] CRAN (R 3.5.1)                  

@@

#> [1] /home/mturgeon/Rlibs

@@

\begin{CodeOutput}
#> Local:    review-max /home/mturgeon/Documents/git_repositories/cbpaper
#> Head:     [f601653] 2019-12-08: Merge branch 'review-max' of https://github.com/sahirbhatnagar/cbpaper into review-max
\end{CodeOutput}
\end{CodeChunk}
